---
title: "Assignmnet 1"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Problem 1.1
Part 1: The F_2(x) cdf will have a heavier lower tail since the cdf is greater, there is more of a probability toward the lower end.

Part2: The F_1(x) cdf will have a heavier upper tail because the cdf is smaller. This means that there will be more of a probability toward the upper end. 

Part 3: F_2(x) will have a larger risk because VaR_0.01 means the 100pth percentile of the loss
distribution, which in this case is the 1 percentile. So when looking at the VaR_0.01 for each cdf, F_2(x) has a higher percentile which tells us it has a higher probability of loss. 


```{r}
## Problem 2
## Part 1
r = 0.2
N = seq(0,40,length=1024)
X = rexp(N, rate = r)

## Part 2
X_theory = dexp(N,rate = r)#r*exp(-r*N) 
#plot(N,X)


hist(X, breaks = 100,prob = TRUE)
lines(N, X_theory, col="red", lty=1)

## Part 3
names = c("gaussian", "rectangular", "triangular", "epanechnikov", "biweight", "cosine", "optcosine")
weights = c(0.1, 1, 2, 5, 10, 50)
for (i in names){
  for(j in weights){
    plot(density(X_theory, bw = j  ,kernel = c(i)), main = "My Kernel Density Plot", xlab = "X-Values", ylab = "Density of my X-Values")
    lines(N, X_theory, col="red", lty=1)
  }
}
## Part 4

## 


```
```{r}
names = c("gaussian")
bandw = c(8,8.5,8.6,8.7,8.8,8.9)#c(1800,1825,1850,1875,1900)

for (i in names){
  for(j in bandw){
    plot(density(X_theory, bw = j  ,kernel = c(i)),main = "My Kernel Density Plot", xlab = "X-Values", ylab = "Density of my X-Values")
    lines(N, X_theory, col="red", lty=1)
  }
}

```

```{r}
## Problem 1.4
## Part 2.1
num = 100
x1 = rlnorm(num,0,1)
y1 = rcauchy(num,0,1)

qqplot(x1,y1,xlab="Log-Norm Dist.", ylab="Cauchy Dist.")

## Part 2.2
x2 = rlnorm(num,0,1)
y2 = rnorm(num,0,1)

qqplot(x2,y2,xlab="Log-Norm Dist.", ylab="Norm Dist.")

```


```{r}
## Problem 1.8
## Part 1
N2 = seq(128)
x_norm = rnorm(N2,0,1)
qqnorm(x_norm, main = "Normal Q-Q Plot",xlab = "Theoretical Quantiles", ylab = "Sample Quantiles")
qqline(x_norm, col = "steelblue", lwd = 2)


## The features that suggest it is a Gaussion distribution are that ........


## Part 2
x_exp = rexp(N2, rate = 1)
qqnorm(x_exp, main = "Normal Q-Q Plot",xlab = "Theoretical Quantiles", ylab = "Sample Quantiles")
qqline(x_exp, col = "steelblue", lwd = 2)
```

```{r}
N = seq(-10,10,length = 1000)
y1 = pcauchy(N,0,1)
plot(N,y1,type="l",col="red")
lines(N, 5*y1, col = 'blue')

qqplot(y1, 5*y1)
```


