---
title: "Fin Stat - Hw2"
author: "Matthew Donaldson"
date: "1/30/2022"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo =TRUE)
```

## Problem 2.2
**1)**
1.1 $F_{m = 0,\lambda = 1,\xi}(x) = 1 - (1 + \xi x)^{-(1/\xi)}$, if $x > 0$; else 0$ This is th cdf from the notes

1.2 The quntile function is the inverse of $F_{m = 0,\lambda = 1,\xi}(x)$ ($Q(u) = F^{-1}(x)$). To derive this I solved for x interms of the variable u. $u = 1 - (1 + \xi x)^{-1/\xi} \rightarrow (1-u)^{-\xi} = 1 + \xi x \rightarrow x = \frac{(1-u)^{-\xi} - 1 }{\xi}$

1.3 If I only had a random generator for a uniform distribution on [0,1] I would generate Monte Carlo samples from the distribution of X by taking the inverse of the uniform distribution to get by in R and then plugging those numbers into the pareto distribution. 

## Problem 2.4
**1)**
```{r}
library(Rsafd)
#head(PCS)
PCS.ml = gpd.ml(PCS[,2]) ## Uses MLE to find estimated parameters of generalized pareto dist.
PCS.ml # Printing those values to see
x = seq(from=-.2,to=15,length=5*length(PCS[,2])) # Generating a sequence that is 5 times the size of the data set used.
X_pareto = rpareto(x,m = 0.07, lambda = 0.7095752,xi=.6359470) # Creating a random sample from the pareto dist. with est. parameters found above. This is the monte carlo simulation
plot(x,X_pareto, main = 'Monte Carlo Random Sample')
#plot(PCS)

```
**2)**
Comments on the qqplot below.....
```{r}
qqplot(PCS[,2], X_pareto)

```
**3)**
```{r}
ks.test(PCS[,2], X_pareto, alternative = c("greater"))

```
**4)**
```{r}
## Part 1
DSPLRET.ml = gpd.ml(DSPLRet) ## Uses MLE to find estimated parameters of generalized pareto dist.
DSPLRET.ml # Printing those values to see
x = seq(from=-.2,to=15,length=5*length(DSPLRet)) # Generating a sequence that is 5 times the size of the data set used.
X_pareto_sp = rpareto(x,m = 3.9550825, lambda = 2.0620189,xi = -0.5748398) # Creating a random sample from the pareto dist. with est. parameters found above. This is the monte carlo simulation
plot(x,X_pareto_sp, main = 'Monte Carlo Random Sample')
plot(DSPLRet)

```
```{r}
# Part 2
qqplot(DSPLRet, X_pareto_sp)
```
```{r}
# Part 3
ks.test(DSPLRet, X_pareto_sp, alternative = c("greater"))
```

## Problem 2.5
```{r}
library(Rsafd)
plot(PSPOT)
qqnorm(PSPOT)
#x = seq(from=-.2,to=15,length=5*length(PSPOT))
#X_norm = rnorm(x)
#qqplot(PSPOT,X_norm) ## This shows that the data has heavy tails
#LMOM_P = sample.LMOM(PSPOT)
#gpd.lmom(LMOM_P)
PSPOT.lmom = gpd.lmom(PSPOT) # This fit GPD to data and gets the shape parameter estimates of gpd
PSPOT.lmom # Prtins parameters
#X_pareto_PSPOT = dpareto(x, m = 15.5282260, lambda = 39.5016753,xi = 0.3958178)

#plot(x,X_pareto_PSPOT)
```

## Problem 2.6
**1-2)**
```{r}
# Part 1
head(DSP)
DSPLRet = log(DSP)
head(DSPLRet)

# Part 2
head(MSP)
MSPLRet = log(MSP)
head(MSPLRet)
```
**3)**
COMMENT ON QQPLOT....
```{r}
## Part 3
qqplot(DSPLRet, MSPLRet, main = 'qqplot of MSPLREt and DSPLRet')
```
**4)**
COMMENTING ON Mean and variance for both and are they similar...
```{r}
# Statistics of DSP
mu_DSPLRet = mean(DSPLRet) # Mean
sigma_DSPLRet = sd(DSPLRet)^2 # Variance
mu_DSPLRet
sigma_DSPLRet
## Statistics for MSP
mu_MSPLRet = mean(MSPLRet) # Mean
sigma_MSPLRet = sd(MSPLRet)^2 # Variance
mu_MSPLRet
sigma_MSPLRet
```
**5)**
```{r}
DSPLRet.lmom = gpd.lmom(DSPLRet) ## Uses L-moments to find estimated parameters of generalized pareto dist.
DSPLRet.lmom # Printing those values to see

MSPLRet.lmom = gpd.lmom(MSPLRet) ## Uses L-moments to find estimated parameters of generalized pareto dist.
MSPLRet.lmom # Printing those values to see

```
