---
title: 'FinStat: Hw 7'
author: "Matthew Donaldson"
date: "3/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 5.12
**1)**
As seen in the output below, slope = 10.3582, intercept = 33.9668. This means 
there is an average wait time between eruptions of 33.9668 minutes and for every 
minute that the eruption lasts it adds 10.3582 minutes to the wait time until
the next eruption.
```{r}
GeyserData = Rsafd::GEYSER

GeyserData = GeyserData[order(GeyserData[,1]),]
duration = GeyserData[,1]
timeNext = GeyserData[,2]

model.lm = lm(timeNext~duration)
summary(model.lm)

```

**2)**
I think that the bandwidth of 0.5(red curve) is the best compared to the other bandwidths used in the ksmooth function becuase of how it shows the trend of the data. Although still not as good as the least squares regression curve seen as the black line in the plot below.
```{r}

model.kb1 = ksmooth(duration, timeNext, kernel = c("normal"), bandwidth = 0.5)
model.kb2 = ksmooth(duration, timeNext, kernel = c("normal"), bandwidth = 1)
model.kb3 = ksmooth(duration, timeNext, kernel = c("normal"), bandwidth = 5)
model.kb4 = ksmooth(duration, timeNext, kernel = c("normal"), bandwidth = 10)
model.kb5 = ksmooth(duration, timeNext, kernel = c("normal"), bandwidth = 50)


plot(duration, timeNext)
lines(duration, model.lm$fitted.values)
lines(duration, model.kb1$y,col = 'red')
lines(duration, model.kb2$y,col = 'blue')
lines(duration, model.kb3$y,col = 'green')
lines(duration, model.kb4$y,col = 'orange')
lines(duration, model.kb5$y,col = 'cyan')



```
## Problem 5.13
**1)**
Both models don't capture the curvature in the upward trend seen in the data. But they are not exactly the same because the l1fit is less effected by outliers so a change in the curvature wouldn't effect it as much as the least squares fit. So maybe the least squares fit does a little better job of showing the upward increase because it is above the blue line by the end.
```{r}
library(Rsafd)
shipData = Rsafd::SHIP

time = 1:length(shipData)

model.lm = lm(shipData~time)
model.abs = l1fit(time,shipData)

model.lm$coefficients
model.abs$coefficients

plot(time,shipData)
lines(time,model.lm$fitted.values , col = 'red')
lines(time,fitted(model.abs), col = 'blue')

```
 
**2)**
Degree 4 seems most reasonable because it captures the trend without starting to 
fit to the noise in the data. Degree 2 also seems okay but does not capture the
first curve in the data.
```{r}
plot(time,shipData)
lines(time, fitted(lm(shipData~poly(time,2))), col= 'red')
lines(time, fitted(lm(shipData~poly(time,4))), col= 'green')
lines(time, fitted(lm(shipData~poly(time,6))), col= 'blue')
lines(time, fitted(lm(shipData~poly(time,8))), col= 'purple')


```

**3)**

The smoothed curves change with the value of df by fitting more to the noise in the data as the df gets higher while the lower df capture the trend but not the noise. I think df = 6 is the best fit because of its ability to capture the trend but not the noise.
```{r}

plot(time,shipData, main="Natural Splines")
lines(time,fitted(lm(shipData~splines::ns(time,df=2))), col = 'red')
lines(time,fitted(lm(shipData~splines::ns(time,df=6))), col = 'green')
lines(time,fitted(lm(shipData~splines::ns(time,df=10))), col = 'blue')
lines(time,fitted(lm(shipData~splines::ns(time,df=14))), col = 'purple')
lines(time,fitted(lm(shipData~splines::ns(time,df=16))), col = 'cyan')


```

**4)**
As the bandwidth increases the line becomes smoother and does a worse job of showing the trend as the line becomes flat. In the smaller bandwidth as seen with the red line (b=1), it fits to the individual points, which is the trend + the noise so a bandwidth of 5 or 20 is more suitable because it shows the trend but not the noise.
```{r}

model.k1 = ksmooth(time, shipData, kernel = c("normal"), bandwidth = 1)
model.k2 = ksmooth(time, shipData, kernel = c("normal"), bandwidth = 5)
model.k3 = ksmooth(time, shipData, kernel = c("normal"), bandwidth = 20)
model.k4 = ksmooth(time, shipData, kernel = c("normal"), bandwidth = 50)
model.k5 = ksmooth(time, shipData, kernel = c("normal"), bandwidth = 125)


plot(time, shipData)
lines(ksmooth(time,shipData, "normal", bandwidth=1), col = 'red')
lines(ksmooth(time,shipData, "normal", bandwidth=5), col = 'green')
lines(ksmooth(time,shipData, "normal", bandwidth=20), col = 'cyan')
lines(ksmooth(time,shipData, "normal", bandwidth=50), col = 'blue')
lines(ksmooth(time,shipData, "normal", bandwidth=125), col = 'orange')



```


## Problem 5.15

**1/2)**
the Sum Squared residuals is pretty high, showing that it is not a great fit.
```{r}

data = Rsafd::ROCK
#head(data)

x = data[,1:3]
y = data[,4]


model = ppr(x,y, nterms=3,max.terms=10)
plot(model$gofn)

SSR = model$gof  #sum((model$residuals)^2)
SSR

```

**3)**
Plot of first three additive terms in the projection pursuit regression formula and the last plot is the scatterplot of f the response against the fitted values and the residuals respectively. The ones in red are the residuals vs. y.
```{r}

plot(model)

plot(y,model$fitted.values)
points(y,model$residuals, col = 'red')

```















